---
title: "1 - RNA Analysis - Counts"
author: "David Alberto García Estrada"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following code allows a transcriptomic analysis to be made from RNA sequencing data that was previously cleaned and mapped to a reference genome.

# Packages

First it is necessary to install the packages that we will use throughout the code.
NOTE: Uncomment the following lines to install the packages, then comment them out again.

```{r}
#if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("limma")
#BiocManager::install("edgeR")
#BiocManager::install("Glimma")
#BiocManager::install("ggplot2")
#BiocManager::install("ggdendro")
#BiocManager::install("reshape2")
#BiocManager::install("multiClust")
#BiocManager::install("topGO")
#BiocManager::install("Rgraphviz")
#BiocManager::install("KEGGREST")
#BiocManager::install("clusterProfiler")
#BiocManager::install("enrichplot")
#BiocManager::install("ComplexHeatmap")
#BiocManager::install("EnhancedVolcano")
#install.packages("dendextend")
#install.packages("data.table")
#install.packages("ggfortify")
#install.packages("ggpmisc")
#install.packages("rafalib")
#install.packages("reshape2")
#install.packages("stats")
#install.packages("pheatmap")
#install.packages("grid")
#install.packages("gplots")
#install.packages("factoextra")
#install.packages("VennDiagram")
#install.packages("pvclust")
#install.packages("tidyr")
#install.packages("cols4all")
#install.packages("ggnewscale")
#install.packages("devtools")
#library("devtools")
#remotes::install_github("mtennekes/cols4all")
#install_github("jokergoo/ComplexHeatmap")
```

We load these packages

```{r, error=FALSE, warning=FALSE, message=FALSE}
library("devtools")
library("data.table")
library("ggplot2")
library("ggfortify")
library("ggpmisc")
library("rafalib")
library("limma")
library("edgeR")
library("Glimma")
library("stats")
library("RColorBrewer")
library("ggdendro")
library("reshape2")
library("grid")
library("gridtext")
library("gplots")
library("factoextra")
library("multiClust")
library("dendextend")
library("topGO")
library("Rgraphviz")
library("VennDiagram")
library("cluster")
library("ggpubr")
library("NbClust")
library("dplyr")
library("clValid")
library("fpc")
library("corrplot")
library("pvclust")
library("viridis")
library("clusterProfiler")
library("enrichplot")
library("ggVennDiagram")
library("ggvenn")
library("ComplexHeatmap")
library("tidyr")
library("cols4all") 
library("ggnewscale")
library("KEGGREST")
library("tibble")
library("EnhancedVolcano")
```

# Counts matrix

The counts matrix is generated from the counts results obtained with **HTseq** and the mapping performed with **HISAT2**, in *Sort* mode.

**Important**.
We set the working address

```{r}
# Display current directory
getwd()
# If necessary, change the path where the stored files are located. 
workingDir <- "."
setwd(workingDir)
```

Specify the address where the counts tables are located.

```{r}
Rawdata <- "CountsTable/"
```

Generate a single counts table
We create a variable - list containing all files with the common word "Results_"

```{r}
# Each of the file names with the pattern "Results_" is stored.
countFiles <- list.files(path = Rawdata, pattern = "Results_")
# Create vectors for the for loop
list <- list()
id <- c()
# Each of the file NAMES in the "countFiles" vector is stored in the 
# vector "treat" and is printed in the variable tab the table of each file is read.
# "countFile" and stored there in the vector "treat" is stored in the # vector "list".
# and in turn the values of the table read and stored in "tab" are added.
# Finally in the vector id the names of the unique IDs that are in tab are stored
for (countFile in countFiles) {
  # countFile <- "Results_IPT17.txt"
  treat <- sub("", "", countFile)
  tab <- read.table(paste0(Rawdata, countFile), 
                    header = FALSE, sep = "\t", quote = "", 
                    row.names = 1, comment.char = "")
  list[[treat]] <- tab
  id <- unique(c(id, rownames(tab)))
}
```

The above is stored in a list that we have to pass to a table data.frame

```{r}
# The "table" includes as row names each of the IDs, in this case the genes.
table <- data.frame(row.names = id) 
#Dimensions of the table that tells us how many IDs/genes we have.
dim(table) # We have a total of 12024 genes.
# Each of the "treat" stored in "list", is stored in the vector 
# which is added to the vector "table" at the end of the last column of the "table".
# Finally, the legend that the sample was added is printed.
for (treat in names(list)) {
  temptable <- list[[treat]]
  table[rownames(temptable), treat] <- temptable[, 1]
  print(paste("The sample", treat, "was successfully added"))
}
# there will be IDs in some treats that have no counts and appear as NAs.
# so it is necessary to convert these to zeros (0)
table[is.na(table)] = 0
```

Here are some details of the full table of accounts

```{r}
# We remove the last rows that summarize data
table <- table[-c(12020:12024),]
```

We save the table in a text file that we call *Counts_from_HISATSort-HTseq.txt*.

```{r}
write.table(table, file = paste0("Counts_from_HISATSort-HTseq.txt"), 
            row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
```

Continued on 2 - RNA Analysis - Grouping and Standardization