---
title: "6 - RNA Analysis - Enrichment"
author: "David Alberto García Estrada"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Enrichment

## GO Terms

The terms for enrichment are obtained from the [JGI Trichoderma atroviride](https://mycocosm.jgi.doe.gov/Triat2/Triat2.home.html) page.

```{r}
# Import the updated GO term table into the new genome
# The table containing the GO terms is read.
ID_GO <- readMappings(paste0(OutpathGO,"Trichoderma.atroviride.GO.txt"),
                      sep="\t", IDsep = ";")
terminos <- c("BP", "CC", "MF")
for (t in terminos){
  # t <- "BP"
  # term list
  # Compiles the list of GO terms such that each list is a character vector
  # containing all the identifiers of the genes that map to their respective term
  list <- annFUN.gene2GO(whichOnto = t, feasibleGenes = NULL, gene2GO = ID_GO)
  TermGo <- data.frame()
  for (i in names(list)){ # # In the variable "i" is stored the term GO
    y <- list[[i]] # We obtain the genes that give with that GO term
    ln <- length(list[[i]]) # The number of genes that match that GO term
    x <- data.frame(GO=rep(i,ln),y) # The table is stored in X and has the following structure
    # GOi term | Gene A
    # GOi term | Gen B
    # Term GOi | Gen C
    TermGo <- rbind(TermGo,x) # Join the tables in TermGo with the previous x
  }
  # Builds a map with GO terms (orders the genes with their respective GO term).
  tGo <- buildGOmap(TermGo)
  # Store a vector with GO terms
  Term <- tGo$GO
  # We add to the GO term its DESCRIPTIVE term
  addterm <- go2term(Term)
  # Change the title of the columns
  colnames(addterm) <- c("GO", "Term")
  # We join the table tGo with addterm by the common feature which is "GO".
  # It looks like this: GO | Gene | Term
  Gomap <- left_join(tGo, addterm, by = "GO") 
  assign(paste("GOGeneTerm_", t, sep = ""), Gomap)
  # We keep all the rows with the GO and Gene columns.
  term2gene <- Gomap[,c("GO", "Gene")]
  assign(paste("GOGene_", t, sep = ""), term2gene)
  # We keep all the rows with the GO and Term columns.
  term2name <- Gomap[,c("GO", "Term")]
  assign(paste("GOName_", t, sep = ""), term2name)
}
```

# Extraction of the GO terms for Closters

We create the folders where the data and graphs will be stored

```{r}
DirEnrichClos <- paste0(DirEnrich, "Closters/")
dir.create(DirEnrichClos, showWarnings = FALSE)

DirEnrichBP <- paste0(DirEnrichClos, "BP/")
dir.create(DirEnrichBP, showWarnings = FALSE)
PlotBP <- paste0(DirEnrichBP, "Plots/")
dir.create(PlotBP, showWarnings = FALSE)

DirEnrichCC <- paste0(DirEnrichClos, "CC/")
dir.create(DirEnrichCC, showWarnings = FALSE)
PlotCC <- paste0(DirEnrichCC, "Plots/")
dir.create(PlotCC, showWarnings = FALSE)

DirEnrichMF <- paste0(DirEnrichClos, "MF/")
dir.create(DirEnrichMF, showWarnings = FALSE)
PlotMF <- paste0(DirEnrichMF, "Plots/")
dir.create(PlotMF, showWarnings = FALSE)
```

```{r}
# Lists for the for loop
list_enrchK_BP <- list()
list_enrchK_CC <- list()
list_enrchK_MF <- list()

# NOTE: ClustFile was created in "Transcriptomic analysis - part 5".
# ClustFile was generated from data of FDR 0.05 and log2FC greater than 1
for (i in names(ClustFile)){
  tryCatch({
  # i <- "Closter 1"
  # I remove all the spaces that the vector i may have and store it in that variable
  names_file <- sub(" ","",i)
  # We put the data from the list in the vector data.dge
  data.dge <- ClustFile[[i]]
  dge_id <- rownames(data.dge)
  # Enrichment
  dge.erch_BP <- enricher(dge_id, TERM2GENE = GOGene_BP, TERM2NAME = GOName_BP, 
                          pAdjustMethod = "none", pvalueCutoff = 0.01, 
                          qvalueCutoff = 1, minGSSize = 1)
  dge.erch_CC <- enricher(dge_id, TERM2GENE = GOGene_CC, TERM2NAME = GOName_CC, 
                          pAdjustMethod = "none", pvalueCutoff = 0.01,
                          qvalueCutoff = 1, minGSSize = 1)
  dge.erch_MF <- enricher(dge_id, TERM2GENE = GOGene_MF, TERM2NAME = GOName_MF, 
                          pAdjustMethod = "none", pvalueCutoff = 0.01, 
                          qvalueCutoff = 1, minGSSize = 1)
  # We store in a list
  list_enrchK_BP[[names_file]] <- dge.erch_BP
  list_enrchK_CC[[names_file]] <- dge.erch_CC
  list_enrchK_MF[[names_file]] <- dge.erch_MF
  # Save tables
  write.table(dge.erch_BP, file = paste0(DirEnrichBP, "BP-Enrichement-", names_file, ".txt"), 
              row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
  write.table(dge.erch_CC, file = paste0(DirEnrichCC, "CC-Enrichement-", names_file, ".txt"), 
              row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
  write.table(dge.erch_MF, file = paste0(DirEnrichMF, "MF-Enrichement-", names_file, ".txt"), 
              row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
  }, error=function(e){cat("ERROR :",conditionMessage(e),"\n")}) 
}

# Graph
for (i in  names(ClustFile)){
  # i <- "Closter 1"
  tryCatch({
    i <- sub(" ","", i)
    # Biological Process
    BP <- dotplot(list_enrchK_BP[[i]], showCategory = 10, 
            color = "pvalue", title = paste0("Enrichment Biological Process in ", i), font.size=10)
    svg(paste0(PlotBP, "DotplotBP_", i, ".svg"), width = 10, height = 10)
    plot(BP)
    dev.off()
    # Component Cellular
    CC <- dotplot(list_enrchK_CC[[i]], showCategory = 10, 
            color = "pvalue", title = paste0("Enrichment Component Celullar in ", i), font.size=10)
    svg(paste0(PlotCC, "DotplotCC_", i, ".svg"), width = 10, height = 10)
    plot(CC)
    dev.off()
    # Molecular Function
    MF <- dotplot(list_enrchK_MF[[i]], showCategory = 10, 
            color = "pvalue", title = paste0("Enrichment Molecular Function in ", i), font.size=10)
    svg(paste0(PlotMF, "DotplotMF_", i, ".svg"), width = 10, height = 10)
    plot(MF)
    dev.off()
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")}) 
}
```

## Specific closter

```{r}
# Closter 3 (up), 1 y 7 (down)
dotplot(list_enrchK_BP[["Closter3"]], showCategory = 10, color = "pvalue", 
        title = "Enrichment Biological Process in Closter 3", font.size=10)

dotplot(list_enrchK_BP[["Closter1"]], showCategory = 10, color = "pvalue", 
        title = "Enrichment Biological Process in Closter 9", font.size=10)

dotplot(list_enrchK_BP[["Closter7"]], showCategory = 10, color = "pvalue", 
        title = "Enrichment Biological Process in Closter 7", font.size=10)

C3.1.7 <- list(Closter3 = list_enrchK_BP[["Closter3"]],
               Closter1 = list_enrchK_BP[["Closter1"]],
               Closter7 = list_enrchK_BP[["Closter7"]])
dp <- dotplot(merge_result(C3.1.7), showCategory = 10, color = "pvalue", 
        title = "Enrichment Biological Process in Closter 3, 1 & 7", font.size=10)
svg(paste0(PlotBP, "DotplotBP_3-1-7.svg"), width = 10, height = 10)
plot(dp)
dev.off()
```


# Extraction of the GO terms for Contrast

We create the folders where the data and graphs will be stored

```{r}
DirGO <- paste0(DirEnrich, "Crontast/")
dir.create(DirGO, showWarnings = FALSE)
```

18 contrasts

```{r}
UpDown <- c("up.", "down.", "")
condicion
contrastes <- contrasts[-c(1,4)]
for (h  in condicion){
  # h <- "FDR_0.05"
  DirGOFDR <- paste0(DirGO, h,"/")
  dir.create(DirGOFDR, showWarnings = FALSE)
  for (i in contrastes){
    # i <- "IPT7vsWT7"
    DirGOCont <- paste0(DirGOFDR, i, "/")
    dir.create(DirGOCont, showWarnings = FALSE)
    table_enrchk <- data.frame()
    for (GO in c("BP", "CC", "MF")){
      tryCatch({
      # GO <- "BP"
      list_enrchK <- list()
      for (j in UpDown){
        # j <- "up."
        tryCatch({
          # Directory GO
          DirTerGO <- paste0(DirGOCont, GO, "/")
          dir.create(DirTerGO, showWarnings = FALSE)
          # Directory GO - Termino
          DirTerGOTer <- paste0(DirTerGO, "Termino/")
          dir.create(DirTerGOTer, showWarnings = FALSE)
          # Directory GO - ID
          DirTerGOID <- paste0(DirTerGO, "ID/")
          dir.create(DirTerGOID, showWarnings = FALSE)
          # Directory GO - Plot
          DirTerGOPlot <- paste0(DirTerGO, "Plot/")
          dir.create(DirTerGOPlot, showWarnings = FALSE)
          # Read files
          grupo <- read.table(file = paste0(outpath, h, "/Tablas/", j, "DEG-", i, "_" , h, ".txt"),
                              header = T, row.names = 1, sep = "\t", comment.char="")
          genes <- rownames(grupo)
          ####
          genelist <- factor(as.integer(geneNames %in% genes))
          names(genelist) <- geneNames
          # GO it can be "BP" biological process, "MF" molecular function, "CC" cellular component.
          GOdata <- new("topGOdata", ontology = GO, allGenes = genelist,
                        annot = annFUN.gene2GO, gene2GO = ID_GO)
          resulFis <- runTest(GOdata, algorithm = "classic", statistic = "fisher")
          allRes <- GenTable(GOdata, classic = resulFis, orderBy = "weight",
                             ranksOf = "classic", topNodes = 20)
          # This function displays the number of the most significant nodes (rectangles) in the area.
          showSigOfNodes(GOdata, score(resulFis), firstSigNodes = 10, useInfo = "all")
          # This function will save the above graphic in your output folder.
          printGraph(GOdata, resulFis, firstSigNodes = 10,
                     fn.prefix = paste0(DirTerGOPlot, j, "DEG", i, "_", h),
                     useInfo = "all", pdfSW = T)
          # Now let's save the list of GO terms enriched in the differential genes.
          write.table(allRes, file = paste0(DirTerGOTer,"GO_", GO, "_", j, "DEG", i, "_", h, ".txt"),
                    sep = "\t")
          # Section for retrieving GO tables with IDs
          allGO <- genesInTerm(GOdata)
          # Function to obtain the names of genes within the GO terms
          SAM_ANOTATION <- lapply(allGO, function(x) x[x %in% genes])
          nodos <- allRes[allRes$Significant > 0,1]
          # Gets the gene names of the supplied GO IDs
          GOIDtoGenes <- SAM_ANOTATION[nodos]
          GOgenes <- as.matrix(GOIDtoGenes)
          write.table(GOgenes,
                      file = paste0(DirTerGOID, "ID_GO_", GO, "_", j, "DEG", i, "_", h, ".txt"),
                      sep = "\t")
          ####
          # DotPlot
          # Directory GO - DotPlot
          DirTerGODotPlot <- paste0(DirTerGO, "DotPlot/")
          dir.create(DirTerGODotPlot, showWarnings = FALSE)
          # Directory GO - DotTab
          DirTerGODotTab <- paste0(DirTerGO, "DotTab/")
          dir.create(DirTerGODotTab, showWarnings = FALSE)
          # Enrichment
          dge.erch <- enricher(genes, TERM2GENE = get(paste0("GOGene_", GO)), 
                               TERM2NAME = get(paste0("GOName_", GO)), 
                               pAdjustMethod = "none", pvalueCutoff = 0.01, 
                               qvalueCutoff = 1, minGSSize = 1)
          # Save
          write.table(dge.erch, 
                      file=paste0(DirTerGODotTab, GO, "_", j, "DEG", i, "_", h, ".txt"), 
                      row.names = TRUE, col.names = NA, quote = FALSE, sep = "\t")
          ####
          if (j == c("")){
            list_enrchK[["All"]] <- dge.erch
            p <- dotplot(list_enrchK[["All"]], showCategory = 15, color = "pvalue",
                         title = paste0(GO, "Enrichment of the ", j, "DEG", i, "_", h), font.size=10)
            svg(paste0(DirTerGODotPlot, "Dotplot_", GO, "_Enrichment_", j, "DEG", i, "_", h, ".svg"),
                width = 15, height = 10)
            plot(p)
            dev.off()
          } else {
            list_enrchK[[j]] <- dge.erch
            p <- dotplot(list_enrchK[[j]], showCategory = 15, color = "pvalue",
                         title = paste0(GO, "Enrichment of the ", j, "DEG", i, "_", h), font.size=10)
            svg(paste0(DirTerGODotPlot, "Dotplot_", GO, "_Enrichment_", j, "DEG", i, "_", h, ".svg"),
                width = 15, height = 10)
            plot(p)
            dev.off()
          }
          ####
        }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
      }
      merge <- merge_result(list_enrchK)
      Enrich <- dotplot(merge, showCategory = 10, color = "pvalue",
                        title = paste0(GO, " Enrichment"), font.size=10)
      svg(paste0(DirTerGODotPlot, "Dotplot_All_", GO, "_Enrichment.svg"), width = 10, height = 10)
      plot(Enrich)
      dev.off()
      # Join
      tab.dge.erch <- merge[,c(1,2,3,6,10)]
      tab.dge.erch$GO <- GO
      tab.dge.erch <- tab.dge.erch[tab.dge.erch[,"Cluster"]!="All",]
      names(tab.dge.erch) <- sub("Cluster", "UpDown", names(tab.dge.erch))
      tab.dge.erch[,"UpDown"] <- sub("up.", "up", tab.dge.erch[,"UpDown"])
      tab.dge.erch[,"UpDown"] <- sub("down.", "down", tab.dge.erch[,"UpDown"])
      table_enrchk <- rbind(table_enrchk, tab.dge.erch)
      # Save
      write.table(table_enrchk, 
                  file=paste0(DirGOCont, "DEG", i, "_", h, ".txt"), 
                  row.names = F, col.names = T, quote = FALSE, sep = "\t")
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
    }
  }
}
```

# Another plots

```{r}
for (h  in condicion){
  # h <- "FDR_0.05"
  DirGOFDR <- paste0(DirGO, h,"/")
  for (i in contrastes){
    tryCatch({
      # i <- "IPT7vsWT7"
      DirGOCont <- paste0(DirGOFDR, i, "/")
      table <- read.table(paste0(DirGOCont, "DEG", i, "_", h, ".txt"), 
                          header = TRUE, sep = "\t", comment.char = "")
      table$UpDown <- as.factor(table$UpDown)
      table$GO <- as.factor(table$GO)
      table <- na.omit(table)
      # Order
      table2 <- table[order(table$Count, decreasing = TRUE),]
      table3 <- table2[order(table2$GO, decreasing = TRUE),]
      table3 <- Reduce(rbind, by(table3, table3["GO"], head, n=30))
      # Plot
      plot <- ggplot(data = table3, aes(y = Description, x = Count, fill = UpDown)) +
        geom_bar(position="dodge", stat="identity") +
      theme_classic() +
      labs(y = "Description", x = "Number of genes",
           title = "Enrichment Gene Ontology") +
      theme(legend.position = "right", 
            legend.text = element_text(size=12, face="bold"),
            legend.title = element_text(size = 12, face = "bold"),
            axis.text = element_text(color = "black", size = 12, face = "bold"),
            axis.title = element_text(color = "black", size = 12, face = "bold"),
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
            plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) +
      # facet_wrap(. ~ table3$GO, scales = "free")
      facet_wrap(table3$GO ~ table3$UpDown, scales = "free")
      svg(paste0(DirGOCont, "EnrichmentPlot_", h, "-", i, ".svg"), 
          width = 20, height = 15)
      plot(plot)
      dev.off()
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }
}

```

Continued on 7 - RNA Analysis - BlastKOALA