---
title: "7 - RNA Analysis - BlastKOALA"
author: "David Alberto García Estrada"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BlastKOALA

Create output folder

```{r}
KOALA <- "BlastKOALA/"
dir.create(KOALA, showWarnings = FALSE)
```

List IDs 

## Contrast

A list with the IDs of each one of the conditions will be obtained in order to make a grep in linux command line

Create output folder

```{r}
KoalaCondition <- paste0(KOALA, "Condition/")
dir.create(KoalaCondition, showWarnings = FALSE)
```

```{r setup, warning=FALSE}
for (c in condicion){
  # c <- "FDR_0.05"
  KoalaFDR <- paste0(KoalaCondition, c, "/")
  dir.create(KoalaFDR, showWarnings = FALSE)
  for (k in contrastes){
    # k <- "IPT7vsWT7"
    KoalaCont <- paste0(KoalaFDR, k, "/")
    dir.create(KoalaCont, showWarnings = FALSE)
    for (ud in UpDown){
      # ud <- "up."
      TablaDEG <- read.table(paste0(outpath, c, "/Tablas/", ud, "DEG-", k, "_", c, ".txt"),
                             header = TRUE, sep="\t", comment.char="")
      colnames(TablaDEG) <- gsub("X", "GeneID", colnames(TablaDEG))
      colnames(TablaDEG) <- gsub(paste0("_", c, ".txt"), "", colnames(TablaDEG))
      TablaDEG$GeneID <- gsub("-T1", "", TablaDEG$GeneID) # without -T1
      TatroID <- TablaDEG$GeneID
      write.table(TatroID, 
                  file = paste0(KoalaCont, "ListIDs_",  ud, "DEG-", k, "_", c, ".txt"),
                  row.names=F, quote=FALSE, sep="\t", col.names = F)
    }
  }
}
```

## Cluster

A list with the IDs of each one of the cluster will be obtained in order to make a grep in linux command line

Create output folder

```{r}
KoalaClus <- paste0(KOALA, "Cluster/")
dir.create(KoalaClus, showWarnings = FALSE)
```

```{r setup, warning=FALSE}
# Unique condition is "FDR_0.05"
for (i in names(ClustFile)){
  # i <- "Closter 1"
  TablaIDs <- rownames(ClustFile[[i]])
  TablaIDs <- gsub("NA", NA, TablaIDs)
  TablaIDs <- TablaIDs[!is.na(TablaIDs)]
  TablaIDs <- gsub("-T1", "", TablaIDs) # without -T1
  write.table(TablaIDs, file = paste0(KoalaClus, "ListIDs_",  i, ".txt"),
              row.names=F, quote=FALSE, sep="\t", col.names = F)
}
```

## Venn

A list with the IDs of each one of the VennDiagram will be obtained in order to make a grep in linux command line

The lists are generated in the following code (8 - RNA Analysis - Venn Diagram), but it is necessary to have the output folder ready

```{r}
KoalaVenn <- paste0(KOALA, "Venn/")
dir.create(KoalaVenn, showWarnings = FALSE)
```

Continued on 8 - RNA Analysis - Venn Diagram